SparseIDS: Learning Packet Sampling with Reinforcement Learning

### 2020-01-23 paper_rl/paper.tex
Changed detail in draw.io illustration

\subsection{Classifier}
\label{classifier}
We implemented a three-layer LSTM-based classifier with 128 neurons at each layer. As the input features we use source port, destination port, protocol identifier, packet length, \gls{iat} to the previous packet in the flow, packet direction (i.e. forward or reverse path) and all TCP flags (0 if the flow is not TCP).
We omitted \gls{ttl} values, as they are likely to lead to unwanted prediction behaviour \cite{bachl_walling_2019}.  Among the used features, source port, destination port and protocol identifier are constant over the whole flow while the others vary.
During flow extraction we used the usual 5-tuple flow key, which distinguishes flows based on the protocol they use and their source and destination port and IP address.
We use Z-score normalization and a train/test split of 2:1.

@@

In what follows, we explore two technical details on how decisions are performed, either continuously or gradually.

\begin{figure}
\center
  \includegraphics[width=0.3\textwidth]{img/rnn-sampling.pdf}
  \caption{SparseIDS framework. Three deep networks are shown: an LSTM classifier, An Actor network and a Critic network. The Actor takes actions based on the feedback from the the Critic network and the classifier. The Critic network tracks the Actor actions based on the classifier confidence and finally, the packets selector parses packets based on the decision of the Actor network.}
  \label{rnn:sampling}
\end{figure}

>>>>

\subsection{Classifier}
\label{classifier}
We implemented a three-layer LSTM-based classifier with 128 neurons at each layer. As the input features we use source port, destination port, protocol identifier, packet length, \gls{iat} to the previous packet in the flow, packet direction (i.e. forward or reverse path) and all TCP flags (0 if the flow is not TCP).
We omitted \gls{ttl} values, as they are likely to lead to unwanted prediction behaviour \cite{bachl_walling_2019}.  Among the used features, source port, destination port and protocol identifier are constant over the whole flow while the others vary.
Additionally, we add the number of skipped packets since the last packet as a feature, which can help the classifier because otherwise it might be confused because of missing packets. 
During flow extraction we used the usual 5-tuple flow key, which distinguishes flows based on the protocol they use and their source and destination port and IP address.
We use Z-score normalization and a train/test split of 2:1.

@@

In what follows, we explore two technical details on how decisions are performed, either continuously or descretely.

\begin{figure}
\centering
  \includegraphics[width=\columnwidth]{img/rnn-sampling.pdf}
  \caption{SparseIDS framework. Three deep networks are shown: an LSTM classifier, An Actor network and a Critic network. The Actor takes actions based on the feedback from the the Critic network and the classifier. The Critic network tracks the Actor actions based on the classifier confidence and finally, the packets selector parses packets based on the decision of the Actor network.}
  \label{rnn:sampling}
\end{figure}

\begin{figure}
\centering
  \includegraphics[width=\columnwidth]{neural_network.pdf}
  \caption{An overview of the complete neural network being used in our experiments. Ocher stands for the input, blue for linear layers, red for LSTM layers and purple for the outputs of the neural network. The numbers in parentheses next to each label stand for the width of that layer (i.e.~the number of neurons (except for the inputs and outputs)). Parts of the neural network are shared between $v_\text{classification}$ and $v_\text{sparsity}$ as well as $\mu$ and $\sigma$ as they output quantities of similar magnitude and thus the gradients that are computed when updating the neural network are of the same order of magnitude as well. Our implementation also supports sharing the LSTM between all three neural network. However, as the five outputs can produce losses that have different order of magnitude, larger gradients or larger losses would ``overwrite'' the gradients of the smaller ones.}
  \label{rnn:neuralNetworkArchitecture}
\end{figure}
